```{r setup, include=FALSE}
# this prevents package loading message from appearing in the rendered version of your problem set
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE)
library(ggplot2)
library(tidyverse)
library(openintro)
library(agridat)
library(boot)
library(ggridges)
library(gridExtra)
```
### 1. Fast Food
a)
```{r}
fastfood <- openintro::fastfood
ggplot(fastfood, aes(restaurant, calories)) + 
  geom_boxplot() + 
  coord_flip() + 
  theme_grey(14)
```

Short Answer: From the graph, it's clear that the median values of most restaurants are around 500. For most restaurants in the graph, 50% of the calories data fall within the interval between 250 and 750. For Chick Fil-A, 75% of the calories data is less than 500. If we don't consider outliers, the maximum calorie in the graph is from Burger King and it's approximately 1250 and the minimum calorie is from Diary Queen. There are five restaurants with outliers, among which Mcdonalds has the most and highest outliers. For Taco Bell, Subway, Sonic, Mcdonalds, Diary Queen, and Burger King, data between the first quartile and the median is more concentrated than that between the median and the third quartile, while the other restaurants are the opposite.

b)
```{r}
ggplot(fastfood,aes(calories)) + 
  geom_histogram(color = "blue", fill = "lightblue")+
  facet_wrap(~restaurant) +
  theme_grey(14)
```

Short Answer: One of the insights that was not visible in the boxplots is the mode. Compared with boxplot, it's easier to find the mode of the data from the histogram because diagrammatic representation of histogram gives the outlook of data which allows viewers to compare data and analyze it.

c)
```{r}
new_data <- filter(fastfood,xor(str_detect(fastfood$item,'Crispy'),str_detect(fastfood$item,'Grilled'))==TRUE)
new_data$cooktype <- ifelse(str_detect(new_data$item,'Crispy')==TRUE,'Crispy','Grilled')
ggplot(new_data,aes(x = calories, color = cooktype, fill = cooktype)) + 
  geom_density(alpha = .5) + 
  theme_grey(14)
```

Short Answer: Based on the graph, crispy food has more calories than grilled food. Most grilled food has calories less than 500, while most crispy food has calories more than 500. In addition, crispy food is more likely to have extremely high calories. When randomly choosing a food, it's more likely to choose the food that has calories between 250 and 750.

### 2. Temporal Lobes
a)
```{r}
mtl <- openintro::mtl
# computer the number of class
nclass.Sturges(mtl$asubic)
ceiling(log(length(mtl$asubic),base=2))+1
hist(mtl$asubic, breaks="Sturges", col = "lightblue")
ggplot(mtl,aes(asubic)) + 
  geom_histogram(color = "blue", fill = "lightblue") + 
  theme_grey(14)
```

Short Answer: The default method base R uses to determine the bin number is to use Sturges formula. The calculation is above, and the answer is 7. The default bin number for ggplot is 30.
I think large bin width is better because it can easily show the underlying pattern (frequency distribution) of the data. If the bin width is too small, it only shows the individual data. Also, since there are only 35 data, it's not a good choice to use ggplot with default 30 bin numbers

b)
```{r}
par(mfrow=c(1,2))
h1 <- hist(mtl$age, right = FALSE, col = "lightblue", main = 'histogram with right open')
h2 <- hist(mtl$age, col = "lightblue", main = 'histogram with right closed')
```

c)
```{r}
par(mfrow=c(1,2))
hist(mtl$age, right = FALSE, breaks = seq(45.5,75.5,5), col = "lightblue", main = 'histogram with right open', axes = FALSE)
axis(1, at = seq(45.5,75.5,5))
axis(2, at = 0:8, las = 1)
hist(mtl$age, breaks = seq(45.5,75.5,5), col = "lightblue", main = 'histogram with right closed', axes = FALSE)
axis(1, at = seq(45.5,75.5,5))
axis(2, at = 0:8, las = 1)
```

Short Answer: To make two histograms identical, the strategy used here is to shift the boundaries of each bin to the right by 0.5 unit, so that the data previously not included on the boundaries will be included now. So each bin will contain the same data, and therefore the graphs will be identical.

### 3. Soybeans
a)
```{r}
ggplot(australia.soybean, aes(sample=yield)) + 
  geom_qq(distribution = qnorm) + 
  geom_qq_line(col = "blue") + 
  facet_wrap(~ loc, nrow = 1)
```

Short Answer: Based on the graph, Lawes and Brookstead are both close to a normal distribution, and Lawes appears to be closest to a normal distribution.

b)
```{r}
Brookstead <- filter(australia.soybean,loc=='Brookstead')
Lawes <- filter(australia.soybean,loc=='Lawes')
Nambour <- filter(australia.soybean,loc=='Nambour')
RedlandBay <- filter(australia.soybean,loc=='RedlandBay')

g1 <- ggplot(Brookstead,aes(yield))+
  geom_histogram(aes(y=..density..),color = "black", fill = "white")+
  geom_density(col='blue')+
  stat_function(fun = dnorm, 
                args = list(mean = mean(Brookstead$yield), 
                            sd = sd(Brookstead$yield)),col='red')+
  ggtitle("Brookstead")
g2 <- ggplot(Lawes,aes(yield))+
  geom_histogram(aes(y=..density..),color = "black", fill = "white")+
  geom_density(col='blue')+
  stat_function(fun = dnorm, 
                args = list(mean = mean(Lawes$yield), 
                            sd = sd(Lawes$yield)),col='red')+
  ggtitle("Lawes")
g3 <- ggplot(Nambour,aes(yield))+
  geom_histogram(aes(y=..density..),color = "black", fill = "white")+
  geom_density(col='blue')+
  stat_function(fun = dnorm, 
                args = list(mean = mean(Nambour$yield), 
                            sd = sd(Nambour$yield)),col='red')+
  ggtitle("Nambour")
g4 <- ggplot(RedlandBay,aes(yield))+
  geom_histogram(aes(y=..density..),color = "black", fill = "white")+
  geom_density(col='blue')+
  stat_function(fun = dnorm, 
                args = list(mean = mean(RedlandBay$yield), 
                            sd = sd(RedlandBay$yield)),col='red')+
  ggtitle("RedlandBay")
grid.arrange(g1,g2,g3,g4,nrow=2)
```

Short Answer: Based on the graph, Lawes and Brookstead appear to be normally distributed.

c)
```{r}
B <- shapiro.test(Brookstead$yield)
L <- shapiro.test(Lawes$yield)
N <- shapiro.test(Nambour$yield)
R <- shapiro.test(RedlandBay$yield)
B
L
N
R
```

Short Answer: The null hypothesis of this test is that the population is normally distributed, while the alternative hypothesis is that the population is not normally distributed. If the p-value is less than the chosen alpha level, then the null hypothesis is rejected and there is evidence that the data tested are not normally distributed. If the p-value is greater than the chosen alpha level, then the null hypothesis can not be rejected.
We assume that the alpha is 0.05 here. For test on Brookstead, the p-value(0.2594) is greater than the alpha(0.05), so we cannot reject the null hypothesis and conclude that yield of Brookstead is normally distributed. 
For test on Lawes, the p-value(0.3279) is greater than the alpha(0.05), so we cannot reject the null hypothesis and conclude that yield of Lawes is normally distributed.
For test on Nambour, the p-value(0.01911) is less than the alpha(0.05), so we reject the null hypothesis and there is evidence that the yield of Nambour is not normally distributed.
For test on RedlandBay, the p-value(0.0004117) is less than the alpha(0.05), so we reject the null hypothesis and there is evidence that the yield of RedlandBay is not normally distributed.

d)

Short Answer: The three methods used in a,b and c almost give the same results. In part(a), we used the qqplot with theoretical normal lines, and from the graph we conclude that Lawes and Brookstead are both close to a normal distribution, and Lawes appears to be closest to a normal distribution. In part(b), we plotted density histograms with density curves and theoretical normal curves and the graph shows that Lawes and Brookstead appear to be normally distributed. In part(c), we performed the Shapiro-Wilk tests and the test results show that Lawes and Brookstead are normally distributed. 

### 4. Doctors
```{r}
smoke <- filter(breslow,smoke==1)
nsmoke <- filter(breslow,smoke==0)
g1 <- ggplot(smoke,aes(age,y))+geom_col(color = "blue", fill = "lightblue")+
  ggtitle("for smoke")
g2 <- ggplot(nsmoke,aes(age,y))+geom_col(color = "blue", fill = "lightblue")+
  scale_y_continuous(limits=c(0, 200))+
  ggtitle("for not smoke")
grid.arrange(g1,g2,nrow=1)
```

### 5. Loans
a)
```{r}
str(loans_full_schema$loan_amount)
summary(loans_full_schema$loan_amount)
sd(loans_full_schema$loan_amount)
ggplot(loans_full_schema,aes(loan_amount))+
  geom_boxplot()
ggplot(loans_full_schema,aes(loan_amount))+
  geom_histogram(color = "blue", fill = "lightblue")
ggplot(loans_full_schema, aes(loan_amount))+
  geom_histogram(aes(y=..density..),color = "black", fill = "white")+
  geom_density(col='blue')+
  stat_function(fun = dnorm, 
                args = list(mean = mean(loans_full_schema$loan_amount), 
                            sd = sd(loans_full_schema$loan_amount)),col='red')
ggplot(loans_full_schema, aes(sample=loan_amount)) + 
  geom_qq(distribution = qnorm) + 
  geom_qq_line(col = "blue")
```

Short Answer: From the histogram, we can observe that it is a skewed right distribution. Most of the observations are small, and a few observations are much larger than other observations. There are several peaks. The mean is 16362 and the median is 14500. The min is 1000, and the max is 40000, so the range is 39000, and the middle 50% of the data lies between 8000 and 24000, so the IQR is 16000. The standard deviation is 10301.96. There are no outliers. Whether it is from the perspective of density curve or qqplot, the data is not normally distributed.

b)
```{r}
ggplot(loans_full_schema,aes(x=loan_amount,y=loan_purpose))+
  geom_boxplot()
```

c)
```{r}
ggplot(loans_full_schema,aes(x=loan_amount,y=loan_purpose))+
  geom_density_ridges(color = "blue", fill = "lightblue") +
  theme_ridges() 
```

d)
Short Answer: I think boxplot is more effective for this dataset because it gives a clear summary of the data such as median, quartiles, range, and outliers. It also provides some indication of the data’s symmetry and skewness. It's easier to visualize the distribution of loan amount and compare the data for each loan purpose. Using ridgeline plot is not a good choice for this dataset because overlap in the plot may obscure underlying patterns, and the shape of the distribution of loan amount according to each loan purpose are similar.